---
title: "Flowering time in B. rapa key"
output: html_notebook
---


During the summer of 2021, researchers at Davidson College explored whether plant size and temperature influenced the timing of flowering in the annual plant Brassica rapa.  You will be analyzing their data.

Import the data file called Brapa.

```{r}

```

Make sure that Treatment is designated as a factor (this code will tell R to treat it as such).

```{r}
Brapa$Treatment<-as.factor(Brapa$Treatment)
```


Variables:
-HeightFirstFlower: the size of the flowering stalk on the day that plants opened their first flower.
-Treatment: plants were grown under moderate (Control) and elevated (Warm) temperatures.
-FloweringDOY: The timing of flowering, measured as the day of year when flowering occured (a value ranging from 0 to 365).

1) Make a scatterplot of height against flowering time with the color of points representing the treatment plants are in. For the color argument, you can specify col=Brapa$Treatment, which will color code the points for each treatment. It will do this alphabetically. Treatment level "Control" comes first alphabetically, and the default R color "black" comes first alphabetically, so control points will be black, while "Warm" points will be red.

Include a legend using the legend() function that I provide for you, which you should enter after the graph is already made. The first argument is where you describe where the legend goes (I picked a place for you). The second argument is what categories you need labeled. The third argument includes the colors to use (I used R's default first and second colors).

```{r}


legend("topright", legend=c("Control", "Warm"), fill=c("black", "red"))
```


2) Are flowering time and height correlated? Let's check! The cor() function takes the two variables as the first two arguments. The third argument is necessary because we have a few NAs in our dataset, and we want R to only include observations where we have two complete values to calculate the correlation.

```{r}
cor(???, ???, use="pairwise.complete.obs")
```


##################################
Simple linear regression 1
##################################
3) Conduct a simple linear regression that assess whether variation in height explains variation in flowering time. In the code below, your response variable goes on the left and your explanatory variable goes on the right. For this code, you can specify the dataset after the "dat" argument, and then just list the column names in the appropriate places in the formula. 

```{r}
mod1<-lm(???~???, dat=???)

summary(mod1)
```

Q) Interpret each of the coefficients and their associated p-values. 

A:


4) Check the assumptions of the linear regression.  Do you think model assumptions have been met?

To extract residuals, you can use the resid() function where you just need the name of the model as the only argument. The plot(mod1) command will produce four graphs - you only want to pay attention to the first one.

```{r}
hist(resid(mod1))
plot(mod1)
```



##################################
Simple linear regression 2
##################################
5) Contrast your first simple linear regression with a new one where you look to see whether treatment explains variation in flowering time. 

```{r}
mod2<-lm(???~???, dat=???)

summary(mod2)
```

Q) Interpret each of the coefficients and their associated p-values. 

A:


6) Check the assumptions of the linear regression.  Do you think model assumptions have been met? In addition to code you used previously, you can use the code below to assess an additional assumption.


```{r}


var.test(resid(mod2)~Brapa$Treatment[which(Brapa$FloweringDOY!="NA")])
#Note that I had to remove the NAs for FloweringDOY to get this to work
```



##################################
Multiple regression
##################################
7) Conduct a multiple regression that assesses whether variation in height explains variation in flowering time similarly in both treatments (that is, including an interaction between height and treatment). In the code below, your response variable goes on the left, and your explanatory variables go on the right (they are multiplied to indicate that we are examining a potential interaction). For this code, you can specify the dataset after the "dat" argument, and then just list the column names in the appropriate places in the formula. 

```{r}
mod3<-lm(???~???*???, dat=???)

summary(mod3)
```

Q) Interpret each of the coefficients and their associated p-values. 

A: 



8) Check the assumptions of the multiple regression.  Do you think model assumptions have been met?

To extract residuals, you can use the resid() function where you just need the name of the model as the only argument. 

```{r}
hist(resid(mod3))
plot(mod3)
var.test(resid(mod3)~Brapa$Treatment[which(Brapa$FloweringDOY!="NA")])
#Note that I had to remove the NAs for FloweringDOY to get this to work
```

A:



9) Add trend lines to the plot you made above. (hint: you have to recopy the plot code below for the trend lines to run)

The abline() function will plot a line for you. The first argument (a=) is where you enter in a value for the intercept. The second argument (b=) is where you enter the value for the slope. For the second line, I also made the line dashed (lty=2) and changed the color.

```{r}
#Code for plot first

#Code for legend next

#Code for lines of best fit
abline(a=coef(mod3)[1], b=coef(mod3)[3])
abline(a=coef(mod3)[1]+coef(mod3)[2], b=coef(mod3)[3]+coef(mod3)[4], lty=2, col="red")
```

