---
title: "Multiple Linear Regression"
---
*adapted from OpenIntro biostats chapter 7 labs

In most practical settings, more than one explanatory variable is likely to be associated with a response. Multiple linear regression is an extension of simple linear regression that allows for more than one predictor variable in a linear model. As with simple linear regression, the response variable must be numerical, but the predictor variables can be either numerical or categorical. 

There are several applications of multiple regression. One of the most common applications in a clinical setting is estimating an association between a response variable and primary predictor of interest while adjusting for possible confounding variables.

First, we will explore multiple regression model by examining the possible association between cognitive function and the use of statins after adjusting for a potential confounder, age.

Reminder: Statins are a class of drug widely used to lower cholesterol. Research suggests that adults with elevated low density lipoprotein (LDL) cholesterol may be at risk for adverse cardiovascular events. A set of guidelines released in 2013 recommended statin therapy in individuals who are at high risk of adverse cardiovascular events, including individuals with Type II diabetes and moderately high LDL and non-diabetic individuals with atherosclerotic cardiovascular disease and high LDL. If these guidelines were to be followed, almost half of Americans ages 40 to 75 and nearly all men over 60 would be prescribed a statin.

However, some physicians have raised the question of whether treatment with a statin might be associated with an increased risk of cognitive decline.

This lab uses data from the Prevention of REnal and Vascular END-stage Disease (PREVEND) study. Clinical and demographic data for 4,095 individuals are stored in the PREVEND dataset in the oibiostat package. 

```{r}
library(oibiostat)
data(prevend.samp)
```



################################
1) Statin use is coded as an integer vector, where 0 represents a non-user and 1 represents a user. Run the code chunk below to convert the variable Statin into a factor variable, with levels NonUser and User. Describe what this code does below the code chunk (look at the data before and after running it).

```{r}
prevend.samp$Statin <- factor(prevend.samp$Statin, levels = c(0,1), labels = c("Nonuser", "User") )
```

A: 




################################
2) Adjust the code in the code chunk below to generate a scatterplot of RFFT score by Age.  Note that R will assign colors alphabetically to levels in alphabetical order.  Black comes before red alphabetically and Nonuser comes before User, therefore black represents Nonusers and red represents users.

```{r}

plot(???, ???,
     pch = 19, col = prevend.samp$Statin, 
     xlab = "Age (yrs)",
     ylab = "RFFT Score",
     main = "")


legend('topright', legend=c("Nonuser", "User"), fill=c("black", "red"), bty='n')

```

Let's look at the correlation between Age and RFFT score.  Interpret the correlation coefficient below.

```{r}
cor(prevend.samp$Age, prevend.samp$RFFT)
```

A: 



################################
3) Let's fit a simple regression model and interpret the slope coefficient. The lm function stands for linear model.  The arguments are Y~X.  Examine the outpout of the model and interpret all of its elements.  

```{r}
statin_lm <- lm(prevend.samp$RFFT~prevend.samp$Statin)
summary(statin_lm)
```

A: 




################################
4) Age is a potential confounder for the relationship between statin use and cognitive function. If older participants tend to use statins, and higher age is associated with lower cognitive ability, perhaps the observed negative association between cognitive ability and statin use is primarily driven by age.

Let's fit a multiple regression model for predicting RFFT score from statin use and age. Examine the output of the model and interpret all of its elements.  

```{r}
model.RFFTvsStatinAge = lm(prevend.samp$RFFT~prevend.samp$Statin + prevend.samp$Age)
summary(model.RFFTvsStatinAge)
```

Interpretation of intercept:

Interpretation of slope for Statin:

Interpretation of slope for Age:

Interpretation of R2:





################################
5) Let's check all assumptions of the linear model. Run the code below and address whether assumptions are met.

```{r}
#Extract residuals
hist(resid(model.RFFTvsStatinAge))

#Examine residuals (here, fitted values = predicted values)
#Just look at the first two plots for now
plot(model.RFFTvsStatinAge)

#Variance test of residuals against Statin use
var.test(resid(model.RFFTvsStatinAge)~prevend.samp$Statin)
```

A:





################################
6) How does the simple linear model compare to the multiple regression?  Answer the questions below:

a) The adjusted R2 value accounts for the increased number of predictors and should be compared across models to determine the best model. How does the adjusted R2 of the simple linear model compare with the multiple linear model? Which do you think is a better model for the data?

A:




b) As in simple linear regression, inferences can be made about the slope parameters estimated by the model slope coefficients. Based on the multiple regression model, is there a statistically significant association between RFFT score and statin use? 

A: 





c) In a clinical setting, the interpretive focus lies on reporting the nature of the association between the primary predictor and the response, while specifying which potential confounders have been adjusted for. Briefly respond to a clinician who is concerned about a possible association between statin use and decreased cognitive function, based on the analyses conducted in this lab.

A: 

        
        
        
################################
7) Let's add the lines of best fit predicted by the models to the graph.

First, let's identify the slopes and intercepts.  The coef function will isolate the coefficients (parameters) from the model.  There are three parameters from this model.  You can list individual parameters by specifying [1], [2], or [3] at the end of the code (I do that in the next code chunk).

```{r}
coef(model.RFFTvsStatinAge)
```


Next, let's add some trend lines.  The abline(a,b) function will let you specify the intercept (a) and the slope (b) for a line.  I wrote the function to graph the line for Statin non-users.  You need to adjust the code to graph the line for Statin users.  Note that you have to run the entire code chunk to see if the lines are working (you can't just run the abline code by itself).

```{r}
##Paste the code for the graph you previously made

plot(prevend.samp$Age, prevend.samp$RFFT,
     pch = 19, col = prevend.samp$Statin, 
     xlab = "Age (yrs)",
     ylab = "RFFT Score",
     main = "")


##line for Statin non-users
abline(a=coef(model.RFFTvsStatinAge)[1], 
       b=coef(model.RFFTvsStatinAge)[3],
       lwd=2)

##line for Statin users
abline(a=coef(model.RFFTvsStatinAge)[1], 
       b=coef(model.RFFTvsStatinAge)[3],
       lwd=2, lty=2, col="red")

```


