---
title: 'Probability: Applied examples'
author: "SMWadgymar"
---


PRACTICE: rbinom, dbinom, pbinom

Do not use ChatGPT to answer these questions - try them on your own based on the information above. For each of the following scenarios, identify which function (rbinom, dbinom, or pbinom) you should use and find the solution.

Scenario 1: You are studying the effect of a toxin on a frog population. It is known that 40% of the frogs exposed to this toxin will die. You want to simulate how many frogs out of a sample of 100 survive after being exposed to the toxin.

```{r}

```


Scenario 2: A predator has a 30% chance of catching its prey during a hunt. You want to find the cumulative probability that the predator will catch 3 or fewer prey in 10 hunts.

```{r}

```

Scenario 3: In a population of 1000 birds, 5% are infected with a virus. You want to simulate the number of infected birds in a random sample of 100 birds.

```{r}

```


Scenario 4: In a population of plants, there is a 10% chance that any given offspring will carry a certain genetic mutation. You have 50 offspring and want to calculate the probability of getting exactly 5 mutants in this population.

```{r}

```


Scenario 5: There is a 20% chance of detecting an endangered species in a specific region during a survey. You want to find the probability that in 15 surveys, the species is detected at least once.

```{r}

```


Scenario 6: You are testing whether light exposure affects the blooming of flowers. You know from previous studies that 70% of flowers bloom when exposed to sunlight. In an experiment with 20 flowers, you want to calculate the probability that 14 flowers will bloom.

```{r}

```


Scenario 7: In a fish population, 15% are infected by a certain parasite. You want to calculate the probability of observing 8 or fewer infected fish in a sample of 50.

```{r}

```


Scenario 8: A bird species has a 60% chance of successfully hatching eggs. You are interested in simulating how many out of 5 eggs in a nest will hatch in each of 20 nests.

```{r}

```


Scenario 9: In a population of insects, 45% of offspring exhibit a specific trait. You want to simulate the number of offspring with this trait in 50 random insect births.

```{r}

```


Scenario 10: You are studying the expression of a gene and know that each gene copy has a 25% chance of being expressed. In a group of 12 gene copies, you want to determine the probability that 3 of them will be expressed.

```{r}

```




Simulation: Why does sample size matter?

Dr. Chen is studying seed germination in wildflowers. She knows that for a particular species, 30% of seeds typically germinate under standard growing conditions.

a) Use rbinom() to simulate 100 experiments, each testing 10 seeds. Save your results.

```{r}

```


b) Create a table of reflecting the proportion of simulations that resulted in zero through ten seeds germinating by modifying the code below as appropriate. Save your results.

```{r}
small_table<-prop.table(table(factor(???, levels=0:10)))
```


c) Calculate theoretical probabilities of having zero through ten seeds germinate using dbinom() (the "true" probabilities, which reflect infinite sample sizes).

```{r}

```

d) Compare your simulation and theoretical probabilities by modifying the code below as appropriate.

```{r}
barplot(rbind(???, ???), beside=T, legend=c("Simulation", "Theoretical"), xlab="Number of germinated seeds")
```

Description of comparison: 


e) Repeat the steps above in a second study where 10 seeds are used in each of 10,000 experiments. How do simulated and theoretical results compare in this case?

```{r}

```

Description of comparison: 
