---
title: "HW4"
output: pdf_document
date: "2024-10-28"
---

Question 1) Six chimpanzees were tested in the length of time it takes to solve a problem. The “before” measurements are minutes for a naïve chimp to solve the problem while the “after” times are minutes taken after training. 

```{r}
Before <- c(6.0, 5.8, 5.7, 5.5, 5.9, 5.2)
After <- c(2.1, 3.6, 2.8, 2.6, 4.2, 2.3)
```

a) Using proper notation, write the appropriate null and alternative hypotheses to determine if chimps solved the puzzle faster after training.



b) Generate the appropriate visualization for these data.

```{r}

```


c) Comment on whether the data appear normal and determine (with justification) what hypothesis test to perform.

```{r}

```


d) Conduct the hypothesis test and interpret the given p-value.

```{r}

```





Question 2) We're going to look at the US Census American Community Survey from 2012. 

```{r}
library(openintro)
data("acs12")
```

Let's explore whether the number of hours people work differs by their highest level of education.

a) State the null and alternative hypotheses appropriate for this study.

A: 


b) Conduct the appropriate statistical test, including all necessary graphs to ensure that conditions for inference have been met to test the hypotheses you stated above. Interpret the P-value. 


```{r}

```

A: 


c) Conduct any necessary post-hoc analyses, if applicable, to make as complete as conclusion as possible and interpret these results. 


```{r}

```

A: 




Question 3) The World Development Indicators datsets are produced the by World Bank Group and include internationally comparable statistics about the development and quality of people's lives across the globe. The 2011 dataset you will work with includes demographic and socioeconomic data from 165 countries. 

a) Load the data.

```{r}
library(oibiostat)
data("wdi.2011")
```

b) Conduct a hypothesis test to determine whether life expectancy differs among countries with low vs. high access to sanitation services. Show all relevant and appropriate work below. Address ALL steps of hypothesis testing.





Question 4) A key part of Pokemon Go is using evolutions to get stronger Pokemon, and a deeper understanding of evolutions is key to being the greatest Pokemon Go player of all time. This data set covers 75 Pokemon evolutions spread across four species.

a) Load the data from the Unit 4 folder. We have an extra line of code here because the dataset was not being loaded as a data frame correctly.

```{r}
pokemon_go<-read.csv("pokemon_go.csv")
```

b) Some of the variables in the dataset are not classified by R correctly (they should be factors and not characters). The code below will remedy that issue.

```{r}
pokemon_go$species<-as.factor(pokemon_go$species)
```


c) Conduct the appropriate hypothesis test to determine whether the post-evolution combat power (cp_new) differs by species. Which species have different average post-evolution combat powers? You can include multiple code chunks to separate the various steps you take to assess or visualize the data. Address ALL steps of hypothesis testing.
