---
title: "Chi-squared"
author: SMWadgymar
---

Goodness of fit - dice

1) Enter the data
```{r}
dice_obs <- c(53222, 52118, 52465, 52338, 52244, 53285)
exp = 52612/315672
dice_exp <- c(exp, exp, exp, exp, exp, exp) 
```

What does each line of code do?:

A:




2) Run and interpret the test
```{r}
chisq.test(dice_obs, p = dice_exp)
```

Interpretation: 

A:




3) If you want to extract the expected values:
```{r}
names(chisq.test(dice_obs, p = dice_exp))
dice_exp_values <- chisq.test(dice_obs, p = dice_exp)$expected
```

What do each of these lines of code do?

A:




4) One way to graph this data (xpd=F keeps bars from running off the graph):
```{r}
barplot(dice_obs, ylim = c(50000, 56000), xpd = FALSE, names=c(1:6), ylab="Number of occurances", xlab="Face of the die") 

abline(h = 52612, lty = 3)
```

What does the abline function do?

A:


What information do you get from this graph?

A:







Test of independence - Popular kids

5) For contingency tables, we need to first create a matrix, then give that to the chi-squared test. Additionally, matrices are well-suited for making barplots of the data. First, make a matrix. The byrow=T argument means you are entering values from left to right.
```{r}
popular <- matrix(c(63, 31, 25, 88, 55, 33, 96, 55, 32), byrow = TRUE, nrow = 3)
```

Look at the matrix. How many rows and columns does it have?

A: 



6) Run and interpret the test.
```{r}
chisq.test(popular)
```

Interpretation: 



7) Graph the result. The beside=T argument keeps the bars next to each other instead of stacked. Note that values will be grouped by the column variable and the legend will represent the row variable
```{r}
barplot(popular, beside = TRUE, names = c("Grades", "Popular", "Sports"), legend = c("4th", "5th", "6th"), ylim = c(0,100))
```

How do the desires of 4th, 5th, and 6th graders compare?

A: 
